Contextuality and the Born rule in fuzzy quantum theories

### 2017-05-01 prop_letter.tex
prop_letter.tex \paragraph{Quantum and Classical IVPMs} Added...

\newpage{}

Quantum IVPMs turns into the classical IVPMs setup~\cite{JamisonLodwick2004}
when the space of events includes only mutually commuting sub-events,
where any subset of $\events$ is called a sub-event space if it satisfies
the follow conditions.

For example, $\events_{0}=\left\{ \mathbb{0},\proj{\ps},\proj{\ms},\mathbb{1}\right\} $
is a commuting sub-event space. Restricting $\bar{\mu}_{1}$ on $\events_{0}$,
$\bar{\mu}_{1}|_{\events_{0}}$, gives a classical IVPM~$\bar{\mu}_{2}$:
\begin{equation}
\begin{aligned}\bar{\mu}_{2}(\mathbb{0}) & =\imposs & \bar{\mu}_{2}(\proj{0}) & =\unknown\\
\bar{\mu}_{2}(\mathbb{1}) & =\necess & \bar{\mu}_{2}(\proj{1}) & =\unknown
\end{aligned}
\end{equation}
If we identify $\proj{0}$ and $\proj{1}$ as the head and the tail
of a coin, respectively, then $\bar{\mu}_{2}$ represents experimenter
know nothing about a coin before she toss it.

\paragraph{Quantum IVPM and States}

Formally, we can ask: what can we deduce about the state of a quantum
system given a quantum interval-valued probability measure, i.e.,
given observations done with finite resources. In the case that the
intervals are infinitely precise for $d\ge3$, the question reduces
to Gleason's theorem which states that the state of the quantum system
is uniquely determined by the probability measure. But surely the
less resources are available, the less precise the intervals, and
the less we expect to know about the state of the system.

To understand the relation between IVPMs and states, it is instructive
to consider the 2-dimensional quantum probability measure~$\mu_{0}:\events\rightarrow[0,1]$
defined as follow:
\begin{equation}
\mu_{0}(P)=\begin{cases}
\mu_{\ket{\ps}}^{\mathrm{B}}(P) & \textrm{, if }P\in\events_{0}\textrm{ ;}\\
\mu_{\ket{0}}^{\mathrm{B}}(P) & \textrm{, otherwis.}
\end{cases}
\end{equation}
$\mu_{0}$ is not induced by the Born rule as a whole, but is consistent
with the Born rule on $\events_{0}$ and some other proper sub-event
spaces. The similar phenomena happens for IVPMs as well. Formally,
given a quantum IVPM~$\bar{\mu}$, the \emph{core} relative to a
sub-event space~$\events'\subseteq\events$ is\footnote{\yutsung{I changed the definition of core to embedding the Born rule.
While this change might make it easier to explain the core of quantum
IVPMs, should we explain the meaning of the Born rule for classical
probability measures? Check this definition in the classical setting,
like commuting observables or product states... etc.}}
\begin{equation}
\coreBorn\left(\bar{\mu},\events'\right)=\set{\rho}{\forall P\in\events'.~\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)}\textrm{ .}
\end{equation}
The core of an interval-valued probability measure is the set of states
it approximates. For example, $\coreBorn\left(\bar{\mu}_{1},\events'\right)$
contains every density matrix. The 2-dimensional measure~$\mu_{0}$
can be understood by core as well. Consider the IVPM~$\bar{\mu}_{0}=\iota_{\infty}\circ\mu_{0}$
whose cores are\begin{subequations}
\begin{eqnarray}
\coreBorn\left(\bar{\mu}_{0},\events_{0}\right) & = & \left\{ \proj{\ps}\right\} \textrm{ ,}\\
\coreBorn\left(\bar{\mu}_{0},\events_{1}\right) & = & \left\{ \proj{0}\right\} \textrm{ ,}\\
\coreBorn\left(\bar{\mu}_{0},\events\right) & = & \emptyset\textrm{ ,}
\end{eqnarray}
\end{subequations}where $\events_{1}=\left\{ \mathbb{0},\proj{0},\proj{1},\mathbb{1}\right\} $.
That is, $\bar{\mu}_{0}$ approximates $\proj{\ps}$ if we are only
allowed to use $\events_{0}$ during the measurement, and approximates
$\proj{0}$ if $\events_{1}$ are the only allowed ones. Since $\bar{\mu}_{0}$
approximates different states on different sub-event space, it intuitively
approximates nothing if every projector is allowed.

Notice that $\bar{\mu}_{0}$ has a non-empty core relative to $\events_{0}$
and $\events_{1}$ which are all commutative. In general, a commutative
sub-event space reduces quantum IVPMs to classical IVPMs. If a classical
IVPM satisfies a coherence condition, convexity, it must approximate
at least one infinitely-precise measure.

\begin{thm}[Shapley~\cite{Shapley1971,Grabisch2016}]\label{thm:Shapley}For
every interval-valued probability measure~$\bar{\mu}:\events'\rightarrow\mathscr{I}$,
if the following conditions are true.
\begin{itemize}
\item $\events'$ is commutative, i.e., $\bar{\mu}$ is classical.
\item $\bar{\mu}$ is convex, i.e., for all commuting $P_{0},P_{1}\in\events$,
$\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$,
where $P_{0}\vee P_{1}=P_{0}+P_{1}-P_{0}P_{1}$~\cite{Griffiths2003}.
\end{itemize}
Then, $\coreBorn\left(\bar{\mu},\events'\right)\ne\emptyset$.\end{thm}

In informal terms, this result states that although measurements done
with limited resources and poor precision may not uniquely identify
the true state of the system, there is always at least one system
that is consistent with the measurements. However, similar to 2-dimensional
$\iota_{\infty}\circ\mu_{0}$, even if a quantum IVPM approximates
some states for each classical sub-event space, it may not approximate
any particular state while considering for all projectors~$\events$.\footnote{\yutsung{It seems that our example in 4 dimensions together with
relativity core would be unnecessary complex. Therefore, I roll our
example back to 3 dimensions.}}

\begin{example}[Three-dimensional quantum three-interval-valued probability
measure]\label{ex:three-dimensional-three-value} Given a three dimensional
Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Let $\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$,
and\begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\textrm{ ,}
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\textrm{ ,}
\end{aligned}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}It is straightforward to verify that $\bar{\mu}$
is a quantum interval-valued probability measure.

@@

Similarly, let $\events_{2}$ be the minimal event space containing
both $\proj{0}$ and $\proj{\ps'}$. We have $\coreBorn\left(\bar{\mu},\events_{2}\right)=\left\{ \proj{1}\right\} $.
Therefore, 
\begin{equation}
\coreBorn\left(\bar{\mu},\events\right)\subseteq\coreBorn\left(\bar{\mu},\events_{1}\right)\cap\coreBorn\left(\bar{\mu},\events_{2}\right)=\emptyset\textrm{ ,}
\end{equation}
and must be empty.\end{example}

We found this kind of incompatibility can be reduced if we increase
the measurement resource, that is, using finer intervals. We conjecture
by consuming more and more measurement resource, it would be more
and more likely that there is an empty core relative to all projectors~$\events$.

\newpage{}

\paragraph{Contextuality Inequality}

Although Cabello proved some inequalities to verify contextuality~\cite{Cabello_2008},
it seems not directly relating to empty/non-empty core. On one hand,
the part using hidden variables cannot really apply on quantum IVPM
expect $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $
case discussed in the Kochen-Specker theorem paragraph. On the other
hand, for the part predicted by quantum mechanics, the product of
observables in inequalities are always either $\mathbb{1}$ or $-\mathbb{1}$
whose the interval-valued probabilities and expectation values are
always $\necess$. In these kind of situation, the inequalities applies
on all IVPMs, no matter whether it has an empty or non-empty core.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{prop}

>>>>

\paragraph{Quantum and Classical IVPMs}

Quantum IVPMs turns into the classical IVPMs setup~\cite{JamisonLodwick2004}
when the space of events includes only mutually commuting sub-events.
For example, $\events_{0}=\left\{ \mathbb{0},\proj{0},\proj{1},\mathbb{1}\right\} $
is a commuting sub-event space. Restricting $\bar{\mu}_{1}$ on $\events_{0}$,
$\bar{\mu}_{1}|_{\events_{0}}$, gives a classical IVPM~$\bar{\mu}_{2}$:
\begin{equation}
\begin{aligned}\bar{\mu}_{2}(\mathbb{0}) & =\imposs & \bar{\mu}_{2}(\proj{0}) & =\unknown\\
\bar{\mu}_{2}(\mathbb{1}) & =\necess & \bar{\mu}_{2}(\proj{1}) & =\unknown
\end{aligned}
\end{equation}
If we identify the density matrices~$\proj{0}$ and $\proj{1}$ as
the head and the tail of a coin, respectively, then $\bar{\mu}_{2}$
represents experimenter know nothing about a coin before she tosses
it, and two classical states~$\proj{0}$ and $\proj{1}$ are both
consistent with $\bar{\mu}_{2}$ on $\events_{0}$.

In general, although measurements done with limited resources and
poor precision may not uniquely identify the true state of a classical
system, there is always at least one system that is consistent with
the measurements.

\begin{thm}[Shapley~\cite{Shapley1971,Grabisch2016}\label{thm:Shapley}]
For every convex IVPM~$\bar{\mu}:\events'\rightarrow\mathscr{I}$,
if $\events'$ commutes, then $\coreBorn\left(\bar{\mu},\events'\right)\ne\emptyset$.
\end{thm}

However, even if a quantum IVPM approximates some states for each
classical sub-event space, it may not approximate any particular state
while considering all projectors~$\events$.

\begin{example}[Three-dimensional quantum three-interval-valued probability
measure\label{ex:three-dimensional-three-value}]
Given a three dimensional Hilbert space with an orthonormal basis
$\left\{ \ket{0},\ket{1},\ket{2}\right\} $. Let $\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$,
and
\begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\textrm{ ,}
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\textrm{ ,}
\end{aligned}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}
It is straightforward to verify that $\bar{\mu}$ is a quantum IVPM.

@@

Similarly, let $\events_{2}$ be the minimal event space containing
both $\proj{0}$ and $\proj{\ps'}$. We have $\coreBorn\left(\bar{\mu},\events_{2}\right)=\left\{ \proj{1}\right\} $.
Therefore, 
\begin{equation}
\coreBorn\left(\bar{\mu},\events\right)\subseteq\coreBorn\left(\bar{\mu},\events_{1}\right)\cap\coreBorn\left(\bar{\mu},\events_{2}\right)=\emptyset\textrm{ ,}
\end{equation}
and must be empty.
\end{example}

We found this kind of incompatibility can be reduced if we increase
the measurement resource, that is, using finer intervals. We conjecture
by consuming more and more measurement resource, it would be more
and more likely that a quantum IVPM has a state consistent with it
on all projectors~$\events$.

\yutsung{Although Cabello proved some inequalities to verify contextuality~\cite{Cabello_2008},
it seems not directly relating to empty/non-empty core. On one hand,
the part using hidden variables cannot really apply on quantum IVPM
expect $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $
case. On the other hand, for the part predicted by quantum mechanics,
the product of observables in inequalities are always either $\mathbb{1}$
or $-\mathbb{1}$ whose the interval-valued probabilities and expectation
values are always $\necess$. In this kind of situation, the inequalities
applies on all IVPMs, no matter whether it has an empty or non-empty
core.}

\bibliography{prop}
